\documentclass[20pt, a4paper]{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{parskip} % http://ctan.org/pkg/parskip
\usepackage[utf8]{inputenc}

\title{Math}
\author{Nick }
\date{November 2020}

\begin{document}

\maketitle

\section{Maxwell's Equation(s)}


How to read this

$e_1$ corresponds to the unit vector pointing in the x direction $\vec i$. $e_2$ corresponds to the unit vector pointing in the y direction $\vec j$, and so on.

$d_1$ is shorthand for the derivative operator $\frac{d}{de_1}$ which is the same as $\frac{d}{dx}$

$e_{23}$ is shorthand for $e_2e_3$

$i$ is defined as the trivector $e_{123}

We define the derivative 
$$ \nabla = (e_1d_1 + e_2d_2 + e_3d_3) $$
Multiplying by it on the left will give a sum of the divergence and curl.  The curl will be expressed as a multivector.


Applying $\nabla$ in geometric calculus is equivalent to applying the div and curl.  So you'll get back a sum of the div and curl but the curl will use bivectors.  This can be demonstrated by just multiplying out the following example.

$$ \left( e_1d_1 +e_2d_2 +e_3d_3 \right) \left(a_1e_1 + a_2e_2 + a_3e_3 \right) $$



$ \vec E, \vec B, \vec J$ are regular vectors that represent the electric field the magnetic field, and the current density (current per unit area).





Maxwell's Equations using curl and divergence operators

$$\nabla \cdot \vec E = \rho / \epsilon_0; \qquad \nabla \cdot \vec B = 0 $$
$$ \nabla \times \vec B - \frac {1}{c^2} d_t \vec E = \mu_0 \vec J ; \qquad 
 \nabla \times \vec E + d_t \vec B = 0 $$


Maxwell's Equations in Geometric Calculus

$$
\left( \frac{ d_t }{c} + \nabla \right) 
\left(\vec E + ic \vec B\right)
= \frac{\rho}{\epsilon_0} -c\mu_0 \vec J
$$

or fully expanded

$$ 
\left( \frac{ d_t }{c}  + e_1d_1 +e_2d_2 +e_3d_3 \right)
\left(E_1e_1 +E_2e_2 +E_3e_3 + c\left(B_1e_{23} +B_2e_{31} +b_3e_{12} \right) \right)
= \frac{\rho}{\epsilon_0} -c\mu_0\left(J_1e_1 + J_2e_2 + J_3e_3\right)
$$

Let's try to get the regular equations out of this one :)

Let's discuss the different parts.


The only way to get a scalar out of the LHS is by dotting the derivative with the electric field.  
 
$$  \left( e_1d_1 +e_2d_2 +e_3d_3 \right) \cdot \left(E_1e_1 +E_2e_2 +E_3e_3 \right) $$

The $e_xe_x$ terms cancel out.  This gives the first equation.  

$$ d_1E_1 + d_2E_2 + d_3E_3 = \frac{\rho}{\epsilon_0} $$

Dotting the derivative with the magnetic field term will give a trivector.  There is no trivector on the right so this must equal zero.

$$ c(d_1B_1 + d_2B_2 + d_3B_3)e_{123} = 0 $$

Now there's the curls and the time derivative.  Woo.  Let's try the curl of the electric field first. This is the other parts of the product that we ignored earlier.

$$(d_1E_2 - d_2E_1)e_{12} + (d2E_3 -d_3E_2)e_{23} + (d_3E_1 - d_1E_3)e_{31}$$

Notice the $e_{31}$ term has been reversed in order so that it corresponds to the positive $e_2$ axis.

The time derivative of the magnetic field term

$$ d_t B_1 e_{23} +d_t B_2 e_{31} +d_t B_3e_{12} $$

Since there is no bivector on the right hand side these must sum to zero.

$$(d_1E_2 - d_2E_1)e_{12} + (d2E_3 -d_3E_2)e_{23} + (d_3E_1 - d_1E_3)e_{31} + d_t B_1 e_{23} +d_t B_2 e_{31} +d_t B_3e_{12}  = 0 $$

or 

$$(d_1E_2 - d_2E_1 +d_t B_3)e_{12} + (d2E_3 -d_3E_2 + d_t B_1)e_{23} + (d_3E_1 - d_1E_3 +d_t B_2 )e_{31}  = 0 $$

Phew.  Now the curl of the magnetic field and the time derivative of the electric field

$$ (d_1 B_2 e_{131} + d_2 B_1 e_{223} ) 
+ (d_2 E_3 e_{212} + d_3 E_2 e_{331} )
+ (d_1 E_3 e_{112} + d_3 E_1 e_{323} )
$$

These terms have the opposite signs of the curl of the electric field and work out to be a regular vector.

$$ (-d_1 B_2 + d_2 B_1 ) e_3
+ (- d_2 E_3 + d_3 E_2 ) e_1
+ (d_1 E_3 - d_3 E_1 ) e_2
$$

You know what else would be a regular vector? The time derivative of the electric field.

$$ \frac{d_t}{c}(E_1e_1 + E_2e_2 + E_3e_3) $$

These add 

$$ (-d_1 B_2 + d_2 B_1 ) e_3
+ (- d_2 E_3 + d_3 E_2 ) e_1
+ (d_1 E_3 - d_3 E_1 ) e_2
+ \frac{d_t}{c}(E_1e_1 + E_2e_2 + E_3e_3)
$$

It equals the only vector on the right.

$$ ( d_2 B_1 -d_1 B_2 + \frac{d_t}{c} E_3 ) e_3
+ (d_3 E_2 - d_2 E_3 + \frac{d_t}{c} E_1 ) e_1
+ (d_1 E_3 - d_3 E_1 + \frac{d_t}{c} E_2 ) e_2

= -c\mu_0(J_1 e_1 + J_2 e_2 + J_3 e_3 )
$$

Booyakasha.


\section{Vector Chain Rule}

$$ h(t) = f(\vec g(t)) $$

$$ (x,y,z) = g(t) $$

$$ h'(t) = \nabla \vec f(\vec g(t)) \cdot \vec g'(t) $$

\section{Directional Derivative}

The directional derivative of the function f at position $\vec a$ in direction $\vec v$.

$$ \nabla_{\vec v} f(\vec a) := \lim_{h \to 0} \frac{ f(\vec a + h \vec v ) - f(\vec a ) } {h} $$

Unlike the scalar derivative, the directional derivative will double if the length of $\vec v$ doubles.

\section{Lagrange Error}


Define the lagrange error as the difference between a given function and its taylor series approximation of degree N centered at x=a.

$$E(x):=f(x)-P_{n,a}(x)$$

Because P is an nth degree polynomial the (n+1)th derivative of the error reduces the taylor series to zero.

$$E^{(n+1)}(x)=f^{(n+1)}(x)$$

If this can be bounded

$$|E^{(n+1)}(x)|=|f^{(n+1)}(x)| \leq M$$

then the error can be bounded like this.

$$
|E(x)|\leq \frac{M(x-a)^{(n+1)}}{(n+1)!}
$$

Proof

$$|E^{(n+1)}(x)|=|f^{(n+1)}(x)| \leq M$$

$$
|\int E^{(n+1)}(x)dx| \leq \
\int|E^{(n+1)}(x)|dx \leq \int M dx
$$

$$
|E^{(n)}(x)| \leq Mx+c
$$

$$
|E^{(n)}(a)|=0 \leq Ma+c
$$

The c which gives the smallest bound is $c=-Ma$

$$
|E^{(n)}(x)|\leq M(x-a)
$$

$$
|E^{(n-1)}(x)|\leq \frac{M(x-a)^2}{2}
$$

Integrate until we get the original error function


$$
|E(x)|\leq \frac{M(x-a)^{(n+1)}}{(n+1)!}
$$



\section{Introduction}


Trigonometry

https://www.eeweb.com/tools/trigonometry-laws-and-identities-sheet/

$$\sin(a+b)=\sin a \cos b + \cos a \sin b \\
$$

Chain Rule

$$(f(g(x)))' = f'(g(x))\cdot g'(x)$$

$$\frac{dy}{dt} = \frac{dy}{dx} \cdot \frac{dx}{dt}$$

Parametric Derivatives

To find the derivative of something with respect to x, take its derivative with respect to t and then divide it by the derivative of x with respect to t.

$$y'(x) = \frac{y'(t)}{x'(t)},  \quad = \frac{dy}{dx} = \frac{dy}{dt} \frac{dt}{dx}
 $$

$$y''(x) = \frac{ (y'(t))' }{ x'(t) } = \frac{ y''(t) }{ x'(t) }$$

Abstract Math for Exponentials

Assume $$ \quad f(c)=f(a+b) = f(a)f(b) \quad \forall a,b \in \mathbb{R}  
$$

$$f(c) = f(a+b) \iff c = a + b $$

$$ f(0) = f(0)f(0) => f(0) = 1 \ | \  f(0)=0 $$

$$ f(0) = 0 = f(a)f(-a) => f(a) = 0 \quad \forall a \in R $$

$$  f(0) = 1 = f(a)f(-a) => f(-a) = (f(a))^{-1} $$ 

$$ f(a)=f(nb) = n f(b) $$

Cross multiplication rule

$$(x+a)(x+b) = x^2 + (a+b)x + ab \\
$$

Quadratic Equation Derivation

\begin{equation}
ax^2 + bx + c = \\ 0
a ( x^2 + bx/a ) + c = 0 \\
a ( x^2 + 2 bx/2a + (b/2a)^2  - (b/2a)^2  ) + c = 0\\
a ( x^2 + 2 bx/2a + (b/2a)^2   ) - b^2/4a + c = 0\\
a ( x + b/2a ) ^2 - b^2/4a + c = 0\\
x = -b/2a \pm \sqrt{ b^2/4a^2 - c/a }
\end{equation}

Imaginary Numbers

$$(ai + bj) ( ci + dj ) = (ac + bd) + (ad -bc) ij$$

Geometric Algebra

Orthogonal vectors anticommute

x and y are unit vectors (0,1) and (1,0)

$$(ax+by)(cx+dy) = (ac+bd) + (ad-bc)xy \\
xx=1\\
yx=-xy$$

$$(xy)^2 = xyxy = -xyyx = -1$$

A product of two orthogonal vectors is isomorphic to the imaginary number.

Parabolic Arc

$$x = (v \cos\theta)t \\
y = (v \sin\theta)t + g t^2 / 2 \\

$$

$$y = \frac {x \sin\theta } {\cos\theta } + \frac{g}{2}\bigg(\frac{x}{v\cos\theta}\bigg)^2 \\ 

y = x \tan\theta + \frac{gx^2}{2v^2}(1+\tan^2\theta)$$

which is quadratic in x and in $\tan\theta$ .

$$\frac{g}{2v^2}(1+\tan^2\theta)x^2 + (\tan\theta) x - y = 0$$

$$\bigg(\frac{gx^2}{2v^2}\bigg) \tan^2\theta + (x)\tan\theta + \bigg(\frac{gx^2}{2v^2} -y \bigg) = 0$$

The angle of elevation to reach a position x,y with velocity v and gravitational acceleration g

$$\tan\theta = \frac{-x \pm \sqrt{x^2 - 4\alpha(\alpha-y)} }{2\alpha} , \quad \alpha = \frac{gx^2}{2v^2}$$

Which angle maximizes the horizontal distance?

A 45 degree angle maximizes the distance of a projectile.  

$$\alpha(1+\tan^2\theta)x^2 + \tan(\theta) x -y = 0,\quad \alpha = \frac{g}{2u^2}$$

$$x = \frac{ -\tan\theta \pm \sqrt{\tan^2\theta - 4\alpha(1+\tan^2\theta)y} } {2\alpha( 1+\tan^2\theta )}$$

If $y=0, \alpha=-1$ then

$$x = \frac{ \tan\theta \mp \sqrt{\tan^2\theta} } { 2(1+\tan^2\theta )} = 0, \frac{ \tan\theta } { (1+\tan^2\theta) }$$

Take $$ x=\frac{z}{1+z^2} \\$$

$$ \frac{dx}{dz}=\frac{1-z^2}{(1+z^2)^2}=0 => z=-1,1 => \theta = 45 $$ degrees

$$\frac{dx}{dz}=\frac{1-z^2}{(1+z^2)^2}$$

The distances is optimal when the rate of change
Let $z=\tan\theta; dz = \sec^2\theta  d\theta$.  Take the non-zero value and find where the rate of change is zero for optimisation.

$$x=2\frac{z}{1+z^2}$$

$$\frac{dx}{dz} = 2 \frac{ (1+z^2) - 2z(z) } { (1+z^2)^2 } = 2\frac{1-z^2}{1+z^2}$$

$$\frac{dx}{d\theta} = 2\frac{1-\tan^2\theta}{1+\tan^2\theta} \sec^2\theta = 0
$$

$dx/d\theta = 0$ when $\tan\theta = \pm1$  when $\theta = 45 $ degrees

A 45 degree angle maximizes the horizontal distance.



\end{document}
